---
title: "My Plotly presentation"
author: "Jeroen van der A"
date: "December 2, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
```

## my Plotly presentation

Hi, In my presentation I'm going to show three plotly plots

- A Basic plotly plot
- A little more advanced
- A bit over the top extravagant chart


## The basic example

```{r slide1}
plot(pressure)
```



## A colored boxplot

```{r slide3}
p <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
p
```



## the more extravagant graph

This chart is a 3d rendering of a path. You could use this for visualize the path of an airplane. 
For example airplanes use fixed paths to land at airfields. This visualization could show the standard path and a number of landing
to show how much the acutal paths vary to the official path.


```{r slide2}
x <- c()
y <- c()
z <- c()
c <- c()

for (i in 1:62) {
  r <- 20 * cos(i / 20)
  x <- c(x, r * cos(i))
  y <- c(y, r * sin(i))
  z <- c(z, i)
  c <- c(c, i)
}  

data <- data.frame(x, y, z, c)

p <- plot_ly(data, x = ~x, y = ~y, z = ~z, type = 'scatter3d', mode = 'lines+markers',
        line = list(width = 6, color = ~c, colorscale = 'Viridis'),
        marker = list(size = 3.5, color = ~c, colorscale = 'Greens', cmin = -20, cmax = 50))

p

```